---
title: "Factors that possibly influence suicide"
author: "Sarah Akinkunmi"
date: "11/20/2021"
output: html_document
---

## My analysis on factors that influence suicide dataset in Austria, Brazil, Denmark, Japan, Norway, Russia Federation, South Africa and the United States of America. This data was taken from Kaggle. 

First, I'll load the csv file:

```{r cars}
library(tidyverse)
file <- read_csv("C:/Sarah's Semicolon files/data-science/project-1/master.csv")
file <- rename(file, `suicide (%)` = `suicides/100k pop`)
```

### Then, I'll extract each country into it's own dataframe:

For Austria:

```{r}
au_df <- file%>%.[1787:2168, ]
```

### For Brazil:

```{r}
br_df <- file%>%.[4173:4544, ]
```

### For Denmark:

```{r}
denmark_df <- file%>%.[7419:7682, ]
```

### For Japan:

```{r}
jp_df <- file%>%.[13365:13736, ]
```

### For Norway:

```{r}
nw_df <- file%>%.[17869:18228, ]
```

### For the Russia Federation:

```{r}
rs_df <- file%>%.[20937:21258, ]
```

### For South Africa:

```{r}
sa_df <- file%>%.[23289:23528, ]
```

### For United States of America:

```{r}
us_df <- file%>%.[26849:27220, ]
```


### Next, I'll perform a chi-square test between the age and generation variables in all the countries.

Null hypothesis (h0): There is no relationship between age and generation
Alternative hypothesis (hA): The two variables are dependent on each other


For Austria:
```{r}
chisq.test(au_df$age, au_df$generation)
```
### Since the p-value is less than 0.05 (2.2e-16), we will reject the null hypothesis. We can conclude that the age and generation depend on each other


## For Brazil:

```{r}
chisq.test(br_df$age, br_df$generation)
```
### Since the p-value is less than 0.05 (2.2e-16), we will reject the null hypothesis. We can conclude that the age and generation depend on each other


## For Denmark:

```{r}
chisq.test(denmark_df$age, denmark_df$generation)
```

### Since the p-value is less than 0.05 (2.2e-16), we will reject the null hypothesis. We can conclude that the age and generation depend on each other


## For Japan:

```{r}
chisq.test(jp_df$age, jp_df$generation)
```

### Since the p-value is less than 0.05 (2.2e-16), we will reject the null hypothesis. We can conclude that the age and generation depend on each other


## For Norway:

```{r}
chisq.test(nw_df$age, nw_df$generation)
```

### Since the p-value is less than 0.05 (2.2e-16), we will reject the null hypothesis. We can conclude that the age and generation depend on each other


## For Russia:

```{r}
chisq.test(rs_df$age, rs_df$generation)
```

### Since the p-value is less than 0.05 (2.2e-16), we will reject the null hypothesis. We can conclude that the age and generation depend on each other


## For South Africa:

```{r}
chisq.test(sa_df$age, sa_df$generation)
```

### Since the p-value is less than 0.05 (2.2e-16), we will reject the null hypothesis. We can conclude that the age and generation depend on each other


## For United States of America: 

```{r}
chisq.test(us_df$age, us_df$generation)
```

### Since the p-value is less than 0.05 (2.2e-16), we will reject the null hypothesis. We can conclude that the age and generation depend on each other


### Our chi-square analysis of age and generation are the same in all countries, so it's safe to conclude that we can use either variable but not both for other analysis

```{r}

```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```
