---
title: "Factors that possibly influence suicide"
author: "Sarah Akinkunmi"
date: "11/20/2021"
output:
  html_document: default
  pdf_document: default
---

### My analysis on factors that influence suicide dataset in Austria, Brazil, Denmark, Japan, Norway, Russian   Federation, South Africa and the United States. This data was taken from Kaggle.   

First, I'll load the csv file:

```{r cars}
library(tidyverse)
file <- read_csv("C:/Sarah's Semicolon files/data-science/project-1/master.csv")
file <- rename(file, `suicide (%)` = `suicides/100k pop`)
```

### Then, I'll extract each country into it's own dataframe:  

For Austria:

```{r}
au_df <- file%>%.[1787:2168, ]
```

For Brazil:

```{r}
br_df <- file%>%.[4173:4544, ]
```

For Denmark:

```{r}
denmark_df <- file%>%.[7419:7682, ]
```

For Japan:

```{r}
jp_df <- file%>%.[13365:13736, ]
```

For Norway:

```{r}
nw_df <- file%>%.[17869:18228, ]
```

For the Russian Federation:

```{r}
rs_df <- file%>%.[20937:21258, ]
```

For South Africa:

```{r}
sa_df <- file%>%.[23289:23528, ]
```

For United States:

```{r}
us_df <- file%>%.[26849:27220, ]
```


### Next, I'll perform a chi-square test between the age and generation variables in all the countries.   

Null hypothesis (h~0~): There is no relationship between age and generation  
Alternative hypothesis (h~a~): The two variables are dependent on each other


For Austria:
```{r}
chisq.test(au_df$age, au_df$generation)
```
Since the p-value is less than 0.05 (2.2e^-16^), we will reject the null hypothesis. We can conclude that the age and generation depend on each other

  

For Brazil:

```{r}
chisq.test(br_df$age, br_df$generation)
```
Since the p-value is less than 0.05 (2.2e^-16^), we will reject the null hypothesis. We can conclude that the age and generation depend on each other  

  
  
For Denmark:

```{r}
chisq.test(denmark_df$age, denmark_df$generation)
```
Since the p-value is less than 0.05 (2.2e^-16^), we will reject the null hypothesis. We can conclude that the age and generation depend on each other  


  

For Japan:

```{r}
chisq.test(jp_df$age, jp_df$generation)
```
Since the p-value is less than 0.05 (2.2e^-16^), we will reject the null hypothesis. We can conclude that the age and generation depend on each other  


  
For Norway:

```{r}
chisq.test(nw_df$age, nw_df$generation)
```

Since the p-value is less than 0.05 (2.2e^-16^), we will reject the null hypothesis. We can conclude that the age and generation depend on each other  


  
For the Russian Federation:

```{r}
chisq.test(rs_df$age, rs_df$generation)
```

Since the p-value is less than 0.05 (2.2e^-16^), we will reject the null hypothesis. We can conclude that the age and generation depend on each other  


  
For South Africa:

```{r}
chisq.test(sa_df$age, sa_df$generation)
```

Since the p-value is less than 0.05 (2.2e^-16^), we will reject the null hypothesis. We can conclude that the age and generation depend on each other  


  
For United States: 

```{r}
chisq.test(us_df$age, us_df$generation)
```

Since the p-value is less than 0.05 (2.2e^-16^), we will reject the null hypothesis. We can conclude that the age and generation depend on each other  
  
  
  
Our chi-square analysis of age and generation are the same in all countries, so it's safe to conclude that we can use either variable but not both for other analysis
  
  
  
  
  
### Next, I'll perform a chi-square test between the age and sex variables in all the countries.  

Null hypothesis (h~0~): There is no relationship between age and generation   
Alternative hypothesis (h~a~): The two variables are dependent on each other  


  
For Austria:
```{r}
chisq.test(au_df$age, au_df$sex)
```
Since our p-value is greater than 0.05 (1), we fail to reject the null hypothesis and we conclude that both variables are independent  
  
  
  
For Brazil:
```{r}
chisq.test(br_df$age, br_df$sex)
```
Since our p-value is greater than 0.05 (1), we fail to reject the null hypothesis and we conclude that both variables are independent
  
  
  
For Denmark:
```{r}
chisq.test(denmark_df$age, denmark_df$sex)
```
Since our p-value is greater than 0.05 (1), we fail to reject the null hypothesis and we conclude that both variables are independent  
  
  
  
For Japan:
```{r}
chisq.test(jp_df$age, jp_df$sex)
```
Since our p-value is greater than 0.05 (1), we fail to reject the null hypothesis and we conclude that both variables are independent  
  
  
  
For Norway:
```{r}
chisq.test(nw_df$age, nw_df$sex)
```
Since our p-value is greater than 0.05 (1), we fail to reject the null hypothesis and we conclude that both variables are independent  
  
  
  
For the Russian Federation:
```{r}
chisq.test(rs_df$age, rs_df$sex)
```
Since our p-value is greater than 0.05 (1), we fail to reject the null hypothesis and we conclude that both variables are independent  
  
  
  
For South Africa:  
```{r}
chisq.test(sa_df$age, sa_df$sex)
```
Since our p-value is greater than 0.05 (1), we fail to reject the null hypothesis and we conclude that both variables are independent  
  
  
  
For United States:
```{r}
chisq.test(us_df$age, us_df$sex)
```
Since our p-value is greater than 0.05 (1), we fail to reject the null hypothesis and we conclude that both variables are independent  
  
  
Therefore, we can safely use both categorical variables in any other analysis  



# ANOVA tests  

One-way ANOVA tests between suicide percentage and age, and the suicide percentage and sex  

Null hypothesis (h~0~): There is no difference between the group means of the independent and residual variables   
Alternative hypothesis (h~a~): The is a difference in group means
  
  
  
For Austria:  
  
Suicide (%) and Age  
```{r}
one_way_au <- aov(`suicide (%)`~ age, data = au_df)
summary(one_way_au)
```
Independent variables being tested in the model (‘age’) and the model residuals (‘Residuals’).   

The Df column displays the degrees of freedom  
For the independent variable, Df = 5  
For the residuals, Df = 376  

The F-value column is the test statistic from the F test. This is the mean square of each independent variable divided by the mean square of the residuals. The larger the F value, the more likely it is that the variation caused by the independent variable is real and not due to chance.   

The Pr(>F) column is the p-value of the F-statistic. This shows how likely it is that the F-value calculated from the test would have occurred if the null hypothesis of no difference among group means were true.
The p-value of the age variable is low (p < 0.001), so it appears that the age groups have an impact on the suicide percentage of Austria.  


Suicide (%) and Sex  

```{r}
one_way_au_1 <- aov(`suicide (%)`~ sex, data = au_df)
summary(one_way_au_1)
```
For the independent variable, Df = 1  
For the residuals, Df = 380    


The p-value of the sex variable is low (p < 0.001), so it appears that the sexes have a significance on the suicide percentage of Austria.  



For Brazil:  
Suicide (%) and Age

```{r}
one_way_br <- aov(`suicide (%)`~ age, data = br_df)
summary(one_way_br)
```
For the independent variable, Df = 5  
For the residuals, Df = 366    


The p-value of the age variable is low (p < 0.001), so it appears that the age groups have a high significance on the suicide percentage of Brazil.  

Suicide (%) and Sex

```{r}
one_way_br_1 <- aov(`suicide (%)`~ sex, data = br_df)
summary(one_way_br_1)
```
For the independent variable, Df = 1  
For the residuals, Df = 370    


The p-value of the sex variable is low (p < 0.001), so it appears that the sexes have a high significance on the suicide percentage of Brazil. 



For Denmark:  
Suicide (%) and Age

```{r}
one_way_dm <- aov(`suicide (%)`~ age, data = denmark_df)
summary(one_way_dm)
```
For the independent variable, Df = 5  
For the residuals, Df = 366    


The p-value of the age variable is low (p < 0.001), so it appears that the age groups have a high significance on the suicide percentage of Denmark. 


Suicide (%) and Sex
```{r}
one_way_dm_1 <- aov(`suicide (%)`~ sex, data = denmark_df)
summary(one_way_dm_1)
```
For the independent variable, Df = 1  
For the residuals, Df = 262   


The p-value of the sex variable is low (p < 0.001), so it appears that the sexes have a high significance on the suicide percentage of Denmark. 


For Japan:  
Suicide (%) and Age

```{r}
one_way_jp <- aov(`suicide (%)`~ age, data = jp_df)
summary(one_way_jp)
```
For the independent variable, Df = 5  
For the residuals, Df = 366    


The p-value of the age variable is low (p < 0.001), so it appears that the age groups have a high significance on the suicide percentage of Japan.  


Suicide (%) and Sex
```{r}
one_way_jp_1 <- aov(`suicide (%)`~ sex, data = jp_df)
summary(one_way_jp_1)
```
For the independent variable, Df = 1  
For the residuals, Df = 370   


The p-value of the sex variable is low (p < 0.001), so it appears that the sexes have a high significance on the suicide percentage of Japan. 


For Norway:  
Suicide (%) and Age

```{r}
one_way_nw <- aov(`suicide (%)`~ age, data = nw_df)
summary(one_way_nw)
```
For the independent variable, Df = 5  
For the residuals, Df = 354   


The p-value of the age variable is low (p < 0.001), so it appears that the age group have a high significance on the suicide percentage of Norway.


Suicide (%) and Sex

```{r}
one_way_nw_1 <- aov(`suicide (%)`~ sex, data = nw_df)
summary(one_way_nw_1)
```
For the independent variable, Df = 1  
For the residuals, Df = 358   


The p-value of the sex variable is low (p < 0.001), so it appears that the sexes have a high significance on the suicide percentage of Norway.


For the Russian Federation:  
Suicide (%) and Age
```{r}
one_way_rs <- aov(`suicide (%)`~ age, data = rs_df)
summary(one_way_rs)
```
For the independent variable, Df = 5  
For the residuals, Df = 316  


The p-value of the age variable is low (p < 0.001), so it appears that the age groups have a high significance on the suicide percentage of the Russian Federation.  


Suicide (%) and Sex
```{r}
one_way_rs_1 <- aov(`suicide (%)`~ sex, data = rs_df)
summary(one_way_rs_1)
```
For the independent variable, Df = 1  
For the residuals, Df = 320 


The p-value of the sex variable is low (p < 0.001), so it appears that the sexes have a high significance on the suicide percentage of the Russian Federation.  


For South Africa:  
Suicide (%) and Age
```{r}
one_way_sa <- aov(`suicide (%)`~ age, data = sa_df)
summary(one_way_sa)
```
For the independent variable, Df = 5   
For the residuals, Df = 234 


The p-value of the age variable is low (p < 0.001), so it appears that the age groups have a high significance on the suicide percentage of South Africa.  


Suicide(%) and Sex
```{r}
one_way_sa_1 <- aov(`suicide (%)`~ sex, data = sa_df)
summary(one_way_sa_1)
```
For the independent variable, Df = 1  
For the residuals, Df = 238


The p-value of the sex variable is low (p < 0.001), so it appears that the sexes have a high significance on the suicide percentage of South Africa. 


For United States:  
Suicide (%) and Age
```{r}
one_way_us <- aov(`suicide (%)`~ age, data = us_df)
summary(one_way_us)
```
For the independent variable, Df = 5 
For the residuals, Df = 366


The p-value of the age variable is low (p < 0.001), so it appears that the age groups have a high significance on the suicide percentage of United States.  


Suicide (%) and Sex
```{r}
one_way_us_1 <- aov(`suicide (%)`~ sex, data = us_df)
summary(one_way_us_1)
```
For the independent variable, Df = 1  
For the residuals, Df = 370


The p-value of the sex variable is low (p < 0.001), so it appears that the sexes have a high significance on the suicide percentage of United States.  



## Two-way ANOVA tests with Age and Sex as the independent variables

For Austria:
```{r}
two_way_au <- aov(`suicide (%)`~ sex + age, data = au_df)
summary(two_way_au)
```
For Brazil:
```{r}
two_way_br <- aov(`suicide (%)`~ sex + age, data = br_df)
summary(two_way_br)
```
For Denmark:  
```{r}
two_way_dm <- aov(`suicide (%)`~ sex + age, data = denmark_df)
summary(two_way_dm)
```
For Japan:  
```{r}
two_way_jp <- aov(`suicide (%)`~ sex + age, data = jp_df)
summary(two_way_jp)
```
For Norway: 
```{r}
two_way_nw <- aov(`suicide (%)`~ sex + age, data = nw_df)
summary(two_way_nw)
```
For the Russian Federation:  
```{r}
two_way_rs <- aov(`suicide (%)`~ sex + age, data = rs_df)
summary(two_way_rs)
```
For South Africa:  
```{r}
two_way_sa <- aov(`suicide (%)`~ sex + age, data = sa_df)
summary(two_way_sa)
```
For United States:  
```{r}
two_way_us <- aov(`suicide (%)`~ sex + age, data = us_df)
summary(two_way_us)
```
